Preparing for Potential Closure of European Airspaces due to Re-entering Space Objects

### 2025-06-09 paper.qmd
spell-checking

Our approach builds on the discretisation of the geospatial risk area.
For this purpose we use a hexagon-based tessellation pattern.
Uber's H3 method ensures non-overlapping hexagons (and 12 pentagons), with an equidistant relationship between all neighbours, and enables an efficient indexing of the tessel-cells. 
While this paper focusses on the determination of the risk exposure per cell, the approach would allow to identify adjacent cells to account for the inherent uncertainty and risk appetite of decision-makers.

@@

* constant debris fall speed $s_{d_k} = \qty{145}{mph} = S$ 
  ([we adopt this value from pag 7-8 in @patera2008])
* constant aircraft speed, $s_{a} = C_t$, equal to its cruise speed, under the
  simplification that the majority of the flight time is spent cruising, 
* aircraft area as viewed from the top, $a_T = W_t \times L_t = T_t$, 
  where $W_t$ is the wingspan, and $L_t$ the length for the specific aircraft
  type $t$
* aircraft area as viewed from the front, $a_F = W_t \times H_t = F_t$, 
  where $W_t$ is the wingspan, and $H_t$ the height for the specific aircraft
  type $t$

@@

Debris that re-enter the Earth’s atmosphere via drag-induced decay will fly
circular orbits prior to re-entry even if initially on elliptical
ones.[^circular-orbit]
A characteristic point for the determination of the re-entry position is the *'ascending node'*.
This denotes the intersection point of the object crossing the plane of Earth's equator if moving from the southern hemnisphere to the northern one.
The respective angular position, $\theta = \omega + \nu$, of the re-entry point from the
ascending node, see @sec-orbitals, is uniformly distributed for uncontrolled
re-entries, $\sigma(\theta) = 1 / (2\pi)$.
This is due to uncertainties in atmospheric drag and the gravitational effects of the
Equatorial bulge.
The height of the atmosphere is not a constant across the globe. Next to the gravitational effects, the atmosphere may also expand due to solar activity.
Given these uncertainties, we assume that the impact longitude, $\lambda$, is uniformly distributed.

@@

In order to derive $\rho_t(h)$, we have taken the trajectories of flown flights
in the \acr{NM} Area, resampled at $\qty{30}{s}$ so that the calculation of the
occupancy, i.e. the time spent in the hex, per aircraft type is just a matter
of counting the number of points per cell.
This binning operation is extremely fast to perform in Uber's H3 gridding system
and a real advantage in processing the huge amount of data point representing
the flight trajectories.

@@

The processing of data for 1 day takes less than 30 min from
data extraction to impact probability calculation on a
high-end laptop (MacBook Pro M1 Max).

>>>>

Our approach builds on the discretisation of the geospatial risk area.
For this purpose we use a hexagon-based tessellation pattern.
Uber's H3 method ensures non-overlapping hexagons (and 12 pentagons), with an equidistant relationship between all neighbours, and enables an efficient indexing of the cells. 
While this paper focusses on the determination of the risk exposure per cell, the approach would allow to identify adjacent cells to account for the inherent uncertainty and risk appetite of decision-makers.

@@

* constant debris fall speed $s_{d_k} = \qty{145}{mph} = S$ 
  ([we adopt this value from pages 7-8 in @patera2008])
* constant aircraft speed, $s_{a} = C_t$, equal to its cruise speed, under the
  simplification that the majority of the flight time is spent cruising, 
* aircraft area as viewed from the top, $a_T = W_t \times L_t = T_t$, 
  where $W_t$ is the wingspan, and $L_t$ the length for the specific aircraft
  type $t$
* aircraft area as viewed from the front, $a_F = W_t \times H_t = F_t$, 
  where $W_t$ is the wingspan, and $H_t$ the height for the specific aircraft
  type $t$

@@

Debris that re-enter the Earth’s atmosphere via drag-induced decay will fly
circular orbits prior to re-entry even if initially on elliptical
ones.[^circular-orbit]
A characteristic point for the determination of the re-entry position is the *'ascending node'*.
This denotes the intersection point of the object crossing the plane of Earth's equator if moving from the southern hemisphere to the northern one.
The respective angular position, $\theta = \omega + \nu$, of the re-entry point from the
ascending node, see @sec-orbitals, is uniformly distributed for uncontrolled
re-entries, $\sigma(\theta) = 1 / (2\pi)$.
This is due to uncertainties in atmospheric drag and the gravitational effects of the
Equatorial bulge.
The height of the atmosphere is not a constant across the globe. Next to the gravitational effects, the atmosphere may also expand due to solar activity.
Given these uncertainties, we assume that the impact longitude, $\lambda$, is uniformly distributed.

@@

In order to derive $\rho_t(h)$, we have taken the trajectories of flown flights
in the \acr{NM} Area, resampled at $\qty{30}{s}$ so that the calculation of the
occupancy, i.e. the time spent in the hex, per aircraft type is just a matter
of counting the number of points per cell.
This binning operation is extremely fast to perform in Uber's H3 hierarchical
geospatial indexing system and a real advantage in processing the huge amount of
data point representing the flight trajectories.

@@

The processing of data for 1 day takes less than 30 minutes from
data extraction to impact probability calculation on a
high-end laptop (MacBook Pro M1 Max).
